On centrality of orthogonal $\Kt$

### 2016-04-24 paper.tex
Finished first 4 pages

\newtheorem{lm}{Lemma} \numberwithin{lm}{section} \Crefname{lm}{Lemma}{Lemmas}
\newtheorem{cl}[lm]{Corollary} \Crefname{cl}{Corollary}{Corollaries}
\newtheorem{prop}[lm]{Proposition} \Crefname{prop}{Proposition}{Propositions}
\newtheorem*{tm*}{Theorem}
\newtheorem*{lm*}{Lemma}
\theoremstyle{definition} \newtheorem{df}[lm]{Definition} \Crefname{df}{Definition}{Definitions}
\newtheorem{example}[lm]{Example} \Crefname{example}{Example}{Examples}
\theoremstyle{remark} \newtheorem{rk}[lm]{Remark} \Crefname{rk}{Remark}{Remarks}
\newtheorem*{mtm}{Main Theorem}
\newtheorem*{atm}{Another Presentation}

@@

In \cite{Sus} Suslin showed for $n\geq 3$ that the elementary group $\E(n, R)$ coincides with the subgroup of $\GL(n,\,R)$ generated by matrices
of the form $t(u,\,v)=1+uv^t$ where $u\in\Um(n,\,R)$, $v\in R^n$ and $u^tv=0$. Here $u^t$ stands for the transpose of $u$.
This result clearly implies that $\E(n,\,R)$ is normal inside $\GL(n,\,R)$.
Afterwards, W.~van~der~Kallen developed Suslin's ideas and showed for $n\geq4$ that the Steinberg group $\St(n,\,R)$ is isomorphic to the group presented by generators
$$\{X(u,\,v)\mid u\in\Um(n,\,R),\ v\in R^n,\ u^tv=0\}$$ and the following list of relations (see~\cite{vdK}):
\setcounter{equation}{0} \renewcommand{\theequation}{K\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add1} \\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v').
\end{align}
This presentation clearly implies that $\phi\colon\St(n,\,R)\rightarrow\E(n,\,R)$ is a central extension.
Notice that our notation for the generators slightly differs from that of~\cite{vdK}, e.\,g. our elements $X(u, v)$ correspond to van der Kallen's $(u, v^t)$.

Recall that an ideal $I\trianglelefteq R$ is called a \emph{splitting ideal} if the canonical projection $R \twoheadrightarrow R/I$ splits as a unital ring morphism.
For a splitting ideal $I\trianglelefteq R$ the \emph{relative Steinberg group} $\St(n,\,R,\,I)$ can be defined as the kernel $\Ker\big(\St(n,\,R)\epi\St(n,\,R/I)\big)$.

For general $I$ the relative Steinberg group is no more a subgroup of $\St(n, R)$ but rather is its central extension.
For the purposes of the present text it suffices to consider only relative Steinberg groups corresponding to splitting ideals $I$.
For more information regarding general relative Steinberg groups we refer the reader to \cite[Section~3]{SCh}.

\section{Yet another presentation for $\St(4, R)$.} \label{sec:yap}
The key ingredient in Tulenbaev's proof of the local-global principle is the construction of the homomorphism $T$ which fits into the following commutative diagram.
$$\xymatrix{\St(n,\,R\ltimes XR_a[X],\,XR_a[X]) \ar@{^{(}->}[r] \ar[d] & \St(n, R \ltimes XR_a[X]) \ar[d]_-{\lambda_a^*} \\
             \St(n,\,R_a[X],\,XR_a[X]) \ar@{-->}[ru]_-{T} \ar@{^{(}->}[r] & \St(n, R_a[X])}$$
Here $a$ is arbitrary nonnilpotent element of $R$, $n\geq5$ and $\lambda_a^*$ is the morphism induced by the morphism $\lambda_a\colon R \to R_a$ of principal localisation at $a$.
Notice that $XR_a[X]$ is a splitting ideal for both $R_a[X]$ and $R \ltimes XR_a[X]$.
It is clear that the image of $T$ is contained in the image of the top arrow, therefore this implies the relative Steinberg groups in the left-hand side of the diagram are isomoprhic.
%This $T$ is an inverse to a map induced by localisation, so that constructing $\mathrm T$ may be treated
%as a proof of excision property for Steinberg group in this very particular situation.

@@

In order to construct the arrow $T$ Tulenbaev uses the following presentation for the relative Steinberg group, see~\cite[Proposition~1.6]{Tul}. 
%Assume that $I\trianglelefteq R$ such that a short exact sequence
%\renewcommand{\theequation}{$\dagger$}
%\begin{align} \xymatrix{0\ar@<0.1ex>[r]^{}&I\ar@<0.1ex>[r]^{}&R\ar@<0.5ex>[r]^{\pi}&R/I\ar@<0.5ex>[l]^{\sigma}\ar@<0.1ex>[r]^{}&0} \end{align} splits, i.e., $\pi\sigma=1$.
\begin{prop}\label{prop:TulPres}
For a splitting ideal $I$ the group $\St(n,\,R,\,I)$ is isomorphic to the group given by generators
$$\{X(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}$$ subject to relations
\setcounter{equation}{0}
\renewcommand{\theequation}{T\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w),\\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'), \label{add2} \\
&X(ur+w,\,v)=X(u,\,vr)X(w,\,v)\,\text{ for }r\in R,\ (u,\,w)\in\Um_{n\times2}(R) \label{add3}.
\end{align}
\end{prop}
Here $\Um_{n\times2}(R)$ denotes the set of $n\times2$ \emph{unimodular matrices}, i.\,e. matrices $M$ such that there exists a $2\times n$ matrix $N$ satisfying $NM=\begin{pmatrix}1&0\\0&1\end{pmatrix}$.

@@

Now, we need to construct a homomorphism
$$\mathrm T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X],\,XR_a[X]),$$
i.e., to find elements in the target group subject to Tulenbaev's relations, but because of T3 one can do it only for $n\geq5$, see~\cite[Lemmas~1.2 and~1.3\,c)]{Tul}.

One can not parametrise generators by a pair of arbitrary vectors since it is not even known that $t(u,\,v)\in\E(n,\,R)$ for arbitrary orthogonal $u$ and $v$~\cite{Rao}.

To generalise Tulenbaev's results for $n=4$ we use a presentation with two types of generators: $F(u,\,v)$ with $u$ good and $v$ arbitrary, which are additive by the second component, and $S(u,\,v)$ with $u$ arbitrary and $v$ good, which are additive by the first component. For $u$ and $v$ both good we require that these two types of generators coincide. More precisely,

\begin{df}
for $I\trianglelefteq R$, $n\geq4$ define $\St^*(n,\,R,\,I)$ to be a group given by the set of generators
\begin{multline*}
\{F(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}\,\cup\\
\cup\{S(u,\,v)\mid u\in I^n,\ v\in\E(n,\,R)e_1,\ u^tv=0\}
\end{multline*}
subject to relations
\setcounter{equation}{0}
\renewcommand{\theequation}{R\arabic{equation}}
\begin{align}
&F(u,\,v)F(u,\,w)=F(u,\,v+w),\\
&S(u,\,v)S(w,\,v)=S(u+w,\,v),\\
&F(u,\,v)F(u',\,v')F(u,\,v)\inv=F(t(u,\,v)u',\,t(v,\,u)\inv v'),\\
&F(u,\,va)=S(ua,\,v)\ \forall\,a\in I,\,(u,\,v^t)=(Me_1,\,e_2^tM\inv),\,M\!\in\E(n,\,R).
\end{align}
\end{df}

One could probably wonder why we assumed only one relation on conjugated element, instead of four different relations on generators of each type conjugated by generators of each type. Three ``missing'' relations actualy hold automaticaly.

\begin{lm}
\label{allyouneedisf}
Denote $\varphi\colon\St^*(n,\,R,\,I)\rightarrow\E(n,\,R)$ the natural map, sending $F(u,\,v)\mapsto t(u,\,v)$, $S(u,\,v)\mapsto t(u,\,v)$.
\begin{enumerate}
\item 
$\St^*(n,\,R,\,I)$ is generated by the set of elements
$$
\{F(u,\,va)\mid a\in I,\ (u,\,v^t)=(Me_1,\,e_2^tM\inv),\,M\in\E(n,\,R)\};
$$
\item
for any $g\in\St^*(n,\,R,\,I)$ holds
\setcounter{equation}{2}
\renewcommand{\theequation}{R\arabic{equation}'}
\begin{align}
&gF(u,\,v)g\inv=F(\varphi(g)u,\,\varphi(g\inv)^tv);
\end{align}
\item
for any $g\in\St^*(n,\,R,\,I)$ holds
\setcounter{equation}{2}
\renewcommand{\theequation}{R\arabic{equation}''}
\begin{align}
&gS(u,\,v)g\inv=S(\varphi(g)u,\,\varphi(g\inv)^tv);
\end{align}
\item
there exists a well-defined ``transpose automorphism''
$$
F(u,\,v)^t=S(v,\,u),\quad S(u,\,v)^t=F(v,\,u).
$$
\end{enumerate}
\end{lm}

\begin{proof}
Take an arbitrary generator of the first type $F(u,\,v)$ and take $M\in\E(n,\,R)$ such that $F(u,\,v)=F(Me_1,\,M^*\tilde v)$. Here $M^*$ stands for a contragradient matrix, $M^*=(M\inv)^t$. By R1 we have
$$
F(u,\,v)=\prod\limits_{k\neq1}F(Me_1,\,M^*e_k\tilde v_k)
$$
where $\tilde v_k$ stands for $k$-th coordinate of $\tilde v=\sum e_k\tilde v_k$. Similarly, with the use of R4 we get 
$$
S(u,\,v)=\prod\limits_{k\neq1}S(Ne_k\tilde u_k,\,N^*e_1)=\prod\limits_{k\neq1}F(Ne_k,\,N^*e_1\tilde u_k)
$$
and obtain a). Obviously, b) follows from a). For c) it is sufficient to show that
$$
F(u,\,v)S(u',\,v')F(u,\,v)\inv=S(t(u,\,v)u',\,t(u,\,v)^*v').
$$
Let $S(u',\,v')=\prod F(Ne_k,\,N^*e_1\tilde u_k)$, then
\begin{multline*}
F(u,\,v)S(u',\,v')F(u,\,v)\inv=F(u,\,v)\prod F(Ne_k,\,N^*e_1\tilde u_k)F(u,\,v)\inv=\\
=\prod F(t(u,\,v)Ne_k,\,t(u,\,v)^*N^*e_1\tilde u_k)=\\
=\prod S(t(u,\,v)Ne_k\tilde u_k,\,t(u,\,v)^*N^*e_1)=S(t(u,\,v)u',\,t(u,\,v)^*v').
\end{multline*}
Finally, d) follows from c).
\end{proof}

>>>>

\newtheorem{lm}{Lemma} \numberwithin{lm}{section} \Crefname{lm}{Lemma}{Lemmas}
\newtheorem{cl}[lm]{Corollary} \Crefname{cl}{Corollary}{Corollaries}
\newtheorem{prop}[lm]{Proposition} \Crefname{prop}{Proposition}{Propositions}
\newtheorem*{tm*}{Theorem}
\newtheorem*{lm*}{Lemma}
\theoremstyle{definition} \newtheorem{df}[lm]{Definition} \Crefname{df}{Definition}{Definitions}
\newtheorem{example}[lm]{Example} \Crefname{example}{Example}{Examples}
\theoremstyle{remark} \newtheorem{rk}[lm]{Remark} \Crefname{rk}{Remark}{Remarks}
\newtheorem*{mtm}{Main Theorem}
\newtheorem*{atm}{Another Presentation}
\newlist{lmlist}{enumerate}{1} \setlist[lmlist]{label={\rm(\roman{lmlisti})}, ref=\thelm.(\roman{lmlisti}),noitemsep} \Crefname{lmlisti}{Lemma}{Lemmas}

@@

In \cite{Sus} Suslin showed for $n\geq 3$ that the elementary group $\E(n, R)$ coincides with the subgroup of $\GL(n,\,R)$ generated by matrices
of the form $t(u,\,v)=1+uv^t$ where $u\in\Um(n,\,R)$, $v\in R^n$ and $u^tv=0$. Here $u^t$ stands for the transpose of $u$.
This result clearly implies that $\E(n,\,R)$ is normal inside $\GL(n,\,R)$.
Afterwards, W.~van~der~Kallen developed Suslin's ideas and showed for $n\geq4$ that the Steinberg group $\St(n,\,R)$ is isomorphic to the group presented by generators
$$\{X(u,\,v)\mid u\in\Um(n,\,R),\ v\in R^n,\ u^tv=0\}$$ and the following list of relations (see~\cite{vdK}):
\setcounter{equation}{0} \renewcommand{\theequation}{K\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add1} \\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'). \label{conj1}
\end{align}
This presentation clearly implies that $\phi\colon\St(n,\,R)\rightarrow\E(n,\,R)$ is a central extension.
Notice that our notation for the generators slightly differs from that of~\cite{vdK}, e.\,g. our elements $X(u, v)$ correspond to van der Kallen's $(u, v^t)$.

Recall that an ideal $I\trianglelefteq R$ is called a \emph{splitting ideal} if the canonical projection $R \twoheadrightarrow R/I$ splits as a unital ring morphism.
For a splitting ideal $I\trianglelefteq R$ the \emph{relative Steinberg group} $\St(n,\,R,\,I) = \St(\rA_{n-1}, R, I)$ can be defined as the kernel $\Ker\big(\St(n,\,R)\epi\St(n,\,R/I)\big)$.

For a general $I$ the relative Steinberg group $\St(\Phi, R, I)$ is no more a subgroup of $\St(\Phi, R)$ but rather is its central extension.
For the purposes of the present text it suffices to consider only relative Steinberg groups corresponding to splitting ideals $I$.
For more information regarding general relative Steinberg groups we refer the reader to \cite[Section~3]{SCh}.

\section{Yet another presentation for $\St(4, R)$.} \label{sec:yap}
The key ingredient in Tulenbaev's proof of the local-global principle is the construction of the homomorphism $T$ which fits into the following commutative diagram.
$$\xymatrix{\St(n,\,R\ltimes XR_a[X],\,XR_a[X]) \ar@{^{(}->}[r] \ar[d] & \St(n, R \ltimes XR_a[X]) \ar[d]_-{\lambda_a^*} \\
             \St(n,\,R_a[X],\,XR_a[X]) \ar@{-->}[ru]_-{T} \ar@{^{(}->}[r] & \St(n, R_a[X])}$$
Here $n\geq5$, $a$ is arbitrary nonnilpotent element of $R$ and $\lambda_a^*$ is the morphism induced by the morphism $\lambda_a\colon R \to R_a$ of principal localisation at $a$.
Notice that $XR_a[X]$ is a splitting ideal for both $R_a[X]$ and $R \ltimes XR_a[X]$.
It is clear that the image of $T$ is contained in the image of the top arrow, therefore this implies that the relative Steinberg groups in the left-hand side of the diagram are isomoprhic.
%This $T$ is an inverse to a map induced by localisation, so that constructing $\mathrm T$ may be treated
%as a proof of excision property for Steinberg group in this very particular situation.

@@

In order to construct the arrow $T$ Tulenbaev uses the following presentation for the relative Steinberg group, see~\cite[Proposition~1.6]{Tul}. 
%Assume that $I\trianglelefteq R$ such that a short exact sequence
%\renewcommand{\theequation}{$\dagger$}
%\begin{align} \xymatrix{0\ar@<0.1ex>[r]^{}&I\ar@<0.1ex>[r]^{}&R\ar@<0.5ex>[r]^{\pi}&R/I\ar@<0.5ex>[l]^{\sigma}\ar@<0.1ex>[r]^{}&0} \end{align} splits, i.e., $\pi\sigma=1$.
\begin{prop}\label{prop:TulPres}
For a splitting ideal $I$ the group $\St(n,\,R,\,I)$ is isomorphic to the group given by generators
$$\{X(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}$$ subject to relations
\setcounter{equation}{0}
\renewcommand{\theequation}{T\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add2}\\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'), \label{conj2}  \\
&X(ur+w,\,v)=X(u,\,vr)X(w,\,v)\,\text{ for }r\in R,\ (u,\,w)\in\Um_{n\times2}(R) \label{add3}.
\end{align}
\end{prop}
Here $\Um_{n\times2}(R)$ denotes the set of $n\times2$ \emph{unimodular matrices}, i.\,e. matrices $M$ such that there exists a $2\times n$ matrix $N$ satisfying $NM=\begin{pmatrix}1&0\\0&1\end{pmatrix}$.

@@

The construction of the homomorphism $T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X]])$
amounts to choosing certain elements in the target group $\St(n,\,R\ltimes XR_a[X]])$ and proving that these elements satisfy the relations \eqref{add2}--\eqref{add3}, see~\cite[Lemmas~1.2 and~1.3\,c)]{Tul}.
The main problem with this recipe is that the assumption $n\geq5$ is essential to check that the relation~\eqref{add3} holds.

It is not possible to choose arbitrary orthogonal vectors $u$, $v$ as parameters for the generators because in this situation it is not even known that $t(u,\,v)\in\E(n,\,R)$, see~\cite{Rao}.

To generalise Tulenbaev's results for $n=4$ we will use a more ``symmetric`` presentation with two types of generators: 
$F(u,\,v)$ with $u$ ''good`` and $v$ arbitrary and $S(u,\,v)$ with $u$ arbitrary and $v$ ''good``.
The generators $F(u,\,v)$ will be additive in the second component, while $S(u,\,v)$ will be additive in the first one.
When $u$ and $v$ are both ''good`` we require these two generators to coincide.
More formally, we have the following presentation.

\begin{df}
For $I\trianglelefteq R$ and $n\geq4$ define $\St^*(n,\,R,\,I)$ to be the group with the set of generators
\begin{multline*}
\{F(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}\,\cup\{S(u,\,v)\mid u\in I^n,\ v\in\E(n,\,R)e_1,\ u^tv=0\}
\end{multline*}
subject to relations
\setcounter{equation}{0}
\renewcommand{\theequation}{R\arabic{equation}}
\begin{align}
&F(u,\,v)F(u,\,w)=F(u,\,v+w), \label{add4}\\
&S(u,\,v)S(w,\,v)=S(u+w,\,v), \label{add5}\\
&F(u,\,v)F(u',\,v')F(u,\,v)\inv=F(t(u,\,v)u',\,t(v,\,u)\inv v'), \label{conj3} \\
&F(u,\,va)=S(ua,\,v)\ \forall\,a\in I,\,(u,\,v^t)=(M e_1,\,e_2^t M\inv),\,M\!\in\E(n,\,R) \label{coinc}.
\end{align}
\end{df}
Notice that we only described how a generator $F(u, v)$ acts via conjugation on another generator $F(u',v')$ and omitted three similar relations involving generators $S(u,v)$ of the second type.
The reason for this is the following lemma which asserts that the ``missing'' relations will follow automatically from \eqref{add4}--\eqref{coinc}.
\begin{lm}
\label{allyouneedisf}
Denote by $\phi\colon\St^*(n,\,R,\,I)\rightarrow\E(n,\,R)$ the natural map sending $F(u,\,v)\mapsto t(u,\,v)$ and $S(u,\,v)\mapsto t(u,\,v)$.
Then the following facts are true.
\begin{lmlist}
\item $\St^*(n,\,R,\,I)$ is generated as an abstract group by the set of elements 
      $$\{F(u,\,va)\mid a\in I,\ (u,\,v^t)=(Me_1,\,e_2^tM\inv),\,M\in\E(n,\,R)\};$$
\item for any $g\in\St^*(n,\,R,\,I)$ one has 
      \setcounter{equation}{2} \renewcommand{\theequation}{R\arabic{equation}'}
      \begin{equation} gF(u,\,v)g\inv=F(\phi(g)u,\,\phi(g\inv)^tv); \end{equation}
\item for any $g\in\St^*(n,\,R,\,I)$ one has
       \setcounter{equation}{2} \renewcommand{\theequation}{R\arabic{equation}''}
      \begin{equation} gS(u,\,v)g\inv=S(\phi(g)u,\,\phi(g\inv)^tv); \end{equation}
\item there is a ``transpose automorphism'' defined on $\St^*(n, R, I)$ satisfying
      $$F(u,\,v)^t=S(v,\,u),\quad S(u,\,v)^t=F(v,\,u).$$
\end{lmlist} \end{lm}

\begin{proof}
Let $F(u,\,v)$ be an arbitrary generator of the first type and let $M\in\E(n,\,R)$ be such that $F(u,\,v)=F(Me_1,\,M^*\tilde v)$. 
Here we denote by $M^*$ the contragradient matrix, $M^*=(M\inv)^t$. 
By \eqref{add4} we have $$F(u,\,v)=\prod\limits_{k\neq1}F(Me_1,\,M^*e_k\tilde v_k)$$
where $\tilde v_k$ stands for $k$-th coordinate of $\tilde v=\sum e_i\tilde v_i$.
Applying relations \eqref{add5}, \eqref{coinc} we get
$$S(u,\,v)=\prod\limits_{k\neq1}S(Ne_k\tilde u_k,\,N^*e_1)=\prod\limits_{k\neq1}F(Ne_k,\,N^*e_1\tilde u_k)$$
and, thus, obtain $(\mathrm{i})$. Obviously, $(\mathrm{ii})$ follows from $(\mathrm{i})$. 
To prove $(\mathrm{iii})$ it suffices to show that
$$F(u,\,v)S(u',\,v')F(u,\,v)\inv=S(t(u,\,v)u',\,t(u,\,v)^*v').$$
For $S(u',\,v')=\prod_{k\neq 1} F(Ne_k,\,N^*e_1\tilde u_k)$ we have
\begin{multline*}
F(u,\,v)S(u',\,v')F(u,\,v)\inv=F(u,\,v)\prod F(Ne_k,\,N^*e_1\tilde u_k)F(u,\,v)\inv=\\
=\prod F(t(u,\,v)Ne_k,\,t(u,\,v)^*N^*e_1\tilde u_k)=\\
=\prod S(t(u,\,v)Ne_k\tilde u_k,\,t(u,\,v)^*N^*e_1)=S(t(u,\,v)u',\,t(u,\,v)^*v').
\end{multline*}
Finally, $(\mathrm{iv})$ follows from $(\mathrm{iii})$.
\end{proof}
